\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-09-07 RA-L/experiment.tex
New Reference style

Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.  These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, from \citet{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open source platform or commercially from~\citet{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace. The lights were controlled using an Arduino Uno board connected to an 8-relay shield.  Above  the table, an overhead machine vision system tracks the position of the swarm.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering up to 100 kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

>>>>

Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.  These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, from \citep{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open source platform or commercially from~\citep{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace. The lights were controlled using an Arduino Uno board connected to an 8-relay shield.  Above  the table, an overhead machine vision system tracks the position of the swarm.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering 100 kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

### 2016-09-07 RA-L/intro.tex
Trying to make introduction flow
New Reference style

\section{Introduction}\label{sec:Intro}
Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
\end{align}
The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, \citet{Ou2013} or \citet{qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
  ex vivo to assemble structures in parallel and 
 through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
These applications require control techniques that can reliably exploit large populations despite high under-actuation.

@@

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\cite{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.  By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}. 
A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.

>>>>

\section{Introduction}\label{sec:Intro}
Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
\end{align}
The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
  ex vivo to assemble structures in parallel and 
 through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
These applications require techniques to shape the swarm that can reliably exploit large populations despite high under-actuation.

@@

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.  By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}. 
A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.

### 2016-09-07 RA-L/simulation.tex
moved unused pics
New Reference style

Two simulations were implemented using wall-friction for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.  All code is available online at \citet{Zhao2016mathematica,Zhao2016mathematicaSquare}.

@@

\subsection{Efficient Control of Covariance}
%A set of simulations were conducted to demonstrate the importance of boundary friction.  
Random disturbances impair the performance of Alg. \ref{alg:PosControl2Robots} and Alg. \ref{alg:PosControlNRobots}. Still, we are able to control covariance of the swarm. This section demonstrates simulations of controlling covariance of the swarm.
These simulations use  the 2D physics engine Box2D, by \citet{catto2010box2d}.
 144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig. \ref{fig:SimCovarianceFuncFrictionOpenLoop}.  All robots had  the same initial conditions, but in four tests the boundary friction was $F_f = \{0,1/3 F, 2/3F, F\}$.
 Without friction, covariance has minimal variation.  As friction increases, the covariance can be manipulated to greater degrees.

@@

 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl}. Fig. \ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ axis were controlled from a set of initial conditions.
\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.eps}
\end{center}
\vspace{-1em}
\caption{\label{fig:CovVarControlPlot}
Closed-loop simulation with 144 disc robots and three sets of initial conditions.  The algorithm tracks goal variance and covariance values (green).  The goal covariance switches sign every 30 s.}
\end{figure}

>>>>

Two simulations were implemented using wall-friction for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.  All code is available online at \citep{Zhao2016mathematica,Zhao2016mathematicaSquare}.

@@

\subsection{Efficient Control of Covariance}
%A set of simulations were conducted to demonstrate the importance of boundary friction.  
Random disturbances impair the performance of Alg. \ref{alg:PosControl2Robots} and Alg. \ref{alg:PosControlNRobots}. Still, we are able to control covariance of the swarm. This section demonstrates simulations of controlling covariance of the swarm.
These simulations use  the 2D physics engine Box2D, by \citep{catto2010box2d}.
 144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig. \ref{fig:SimCovarianceFuncFrictionOpenLoop}.  All robots had  the same initial conditions, but in four tests the boundary friction was $F_f = \{0,1/3 F, 2/3F, F\}$.
 Without friction, covariance has minimal variation.  As friction increases, the covariance can be manipulated to greater degrees.

@@

 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl}. Fig. \ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ axis were controlled from a set of initial conditions.
\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:CovVarControlPlot}
Closed-loop simulation with 144 disc robots and three sets of initial conditions.  The algorithm tracks goal variance and covariance values (green).  The goal covariance switches sign every 30 s.}
\end{figure}
